---
title: "Homework 4 - Divya"
author: "Divya Shah"
format: html
---

---
title: "Homework 4 - Divya"
author: "Divya Shah"
format: html
execute:
  warning: false
  message: false
---

## Task 1: Conceptual Questions

### 1. Rewrite the nested function call using base R's chaining operator

```{r}
library(tidyverse)

as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

### 2. Long vs. Wide Format Data

Long format has multiple rows for each observation, with values stored in a single column and another column identifying the variable or category.

Wide format has different variables or measurements spread across separate columns.

Long format is generally preferred for statistical analysis because it works well with many R functions and makes grouping and analysis easier.


## Task 2: Data Manipulation

### Glass Data

First, I read the glass data directly from the URL and add the appropriate column names.

```{r}
glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "Id", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "Type_of_glass"
  ),
  show_col_types = FALSE
)

glass
```

Next, I replace the numeric glass type values with descriptive factor labels. I then keep only observations with an Fe value less than 0.2 and a glass type of tableware or headlamps.

```{r}
glass_filtered <- glass |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_glass %in% c("tableware", "headlamps")
  )

glass_filtered
```

### Yeast Data

I read the yeast data directly from the URL and add the appropriate column names.

```{r}
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  ),
  show_col_types = FALSE
)

yeast
```

I remove the `seq_name` and `nuc` columns. Then I group the data by class and add the mean and median of each numeric variable.

```{r}
yeast_summary <- yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(
    across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(
        mean = ~mean(.x, na.rm = TRUE),
        median = ~median(.x, na.rm = TRUE)
      )
    )
  ) |>
  ungroup()

yeast_summary
```

### Wine Data

First, I read the white wine data from the Excel file located in the main repository folder.

```{r}
library(readxl)

white_wine <- read_excel(
  "../white-wine.xlsx",
  sheet = 1
)

white_names <- read_excel(
  "../white-wine.xlsx",
  sheet = 2
)

colnames(white_wine) <- white_names[[1]]

white_wine <- white_wine |>
  mutate(type = "white")

white_wine
```

Next, I read the red wine data directly from the URL, use the same column names, and identify the observations as red wine.

```{r}
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";",
  col_types = cols(.default = col_double())
)

colnames(red_wine) <- white_names[[1]]

red_wine <- red_wine |>
  mutate(type = "red")

red_wine
```

I combine the white and red wine data.

```{r}
wine_data <- bind_rows(
  white_wine,
  red_wine
)

wine_data
```

Next, I keep only high-quality wines with reasonable alcohol values, sort them from highest to lowest quality, and select the requested variables.

I also calculate the mean and standard deviation of alcohol for each quality level.

```{r}
wine_filtered <- wine_data |>
  filter(
    quality > 6.5,
    alcohol < 13.2
  ) |>
  arrange(desc(quality)) |>
  select(
    contains("acid"),
    alcohol,
    type,
    quality
  ) |>
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol, na.rm = TRUE),
    sd_alcohol = sd(alcohol, na.rm = TRUE)
  ) |>
  ungroup()

wine_filtered
```


## Task 3: Database Practice

First, I connect to the Lahman database.

```{r}
#| message: false
#| warning: false

library(DBI)
library(RSQLite)
library(dplyr)
library(dbplyr)

con <- dbConnect(
  SQLite(),
  "data/Lahman.db"
)
```

### 11. View the Tables

I list all of the tables in the database.

```{r}
dbListTables(con)
```

### 12. Teams from 2015 Using dplyr

I use `tbl()` and dplyr to return all of the data from the Teams table for 2015.

```{r}
teams_2015 <- tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()

teams_2015
```

### 13. Teams from 2015 Using SQL

I repeat the previous question using SQL inside `tbl()`.

```{r}
teams_2015_sql <- tbl(
  con,
  sql("SELECT * FROM Teams WHERE yearID = 2015")
) |>
  collect()

teams_2015_sql
```

### 14. Hall of Fame Players

I return the player ID, year they were voted into the Hall of Fame, and their category. I keep only players who were actually inducted.

```{r}
hall_of_fame <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(
    playerID,
    yearID,
    category
  ) |>
  collect()

hall_of_fame
```

### 15. Add Player Names

I combine the Hall of Fame results with the People table to add each player's first and last name.

```{r}
people_names <- tbl(con, "People") |>
  select(
    playerID,
    nameFirst,
    nameLast
  ) |>
  collect()

hall_of_fame_names <- hall_of_fame |>
  left_join(
    people_names,
    by = "playerID"
  )

hall_of_fame_names
```

### 16. Manager Career Records

I return the player ID, games managed, wins, and losses from the Managers table. I then calculate career totals for each manager.

```{r}
manager_records <- tbl(con, "Managers") |>
  select(
    playerID,
    G,
    W,
    L
  ) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  ) |>
  collect() |>
  mutate(
    win_loss_pct = Total_W / G_managed
  ) |>
  arrange(desc(win_loss_pct))

manager_records
```

### 17. Hall of Fame Managers

Finally, I combine the Hall of Fame information with the manager career records. This keeps only people who were inducted into the Hall of Fame and also managed a team.

```{r}
hall_of_fame_managers <- hall_of_fame_names |>
  inner_join(
    manager_records,
    by = "playerID"
  )

hall_of_fame_managers
```

The final data set should contain 9 columns.

```{r}
ncol(hall_of_fame_managers)
```

Finally, I disconnect from the database.

```{r}
dbDisconnect(con)
```