---
title: "HW 4"
format: html
editor: visual
execute:
  warning: false
---

## Task 1

1.  Rewrite the following nested function call using baseR’s chaining operator:

    ```{r}
    library(tidyverse)
    arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
    ```

    With \|\> pipe operator

    ```{r}
    iris |>
      as_tibble() |>
      select(starts_with("Petal"), Species) |>
      filter(Petal.Length < 1.55) |>
      arrange(Species)
    ```

2.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

    - Wide format data has more than one observation per row (for each x there are multiple ys, i.e. columns), whereas long format stores the same data in multiple rows.

    - Long format might look less pleasing/redundant visually, but is preferred for data analysis (`tidr` can handle/switch between both).

## Task 2

#### Glass Data

3.  Start a chain on your data object that overwrites the Type_of_glass variable using mutate().
    3.  Create a factor version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values). This could also be done by first using an ifelse() type call and then converting to a factor. Either way is fine!
4.  Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of Glass is either “tableware” or “headlamp”.

```{r}
glass_data <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")
)
```

```{r}
glass <- glass_data |>                                     # Q3
  mutate(Type_of_glass = factor(
    Type,
    levels = c(1, 2, 3, 4, 5, 6, 7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
    ))) |>
  filter(Fe < 0.2,                                         # Q4
         Type_of_glass %in% c("tableware", "headlamps"))

head(glass)
```

#### Yeast Data

5.  Read Yeast data; Start a chain that removes the `seq_name` and `nuc` columns.

6.  Continue your chain to add columns corresponding to the mean and median of each numeric variable (`mcg`, `gvh`, `alm`, `mit`, `erl`, `pox`, and `vac`) at each `class` grouping (see the `across()` function as we did in the dplyr video!).

```{r}
yeast_data <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
)
```

```{r}
yeast <- yeast_data |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(mean = mean, median = median)
      )
  )
head(yeast)
```

#### Wine Data

7.  Start a chain on your new combined data object to filter the data to only look at high-quality wines (quality \> 6.5) and wines that have a reasonable alcohol value (alcohol \< 13.2).

8.  Continue your chain to now sort the data from highest quality to lowest

9.  Continue your chain to select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable.

10. Continue your chain to add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

```{r}
library(readxl)

white_wine <- read_excel("../HW3/data/white-wine.xlsx", sheet = 1)
new_names <- read_excel("../HW3/data/white-wine.xlsx", sheet = 2)
colnames(white_wine) <- new_names[[1]]
white_wine$type<-"white"

red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim=";")
colnames(red_wine) <- new_names[[1]]
red_wine$type <- "red"

wine_data <- bind_rows(white_wine, red_wine)
#head(wine_data)
```

```{r}
wine <- wine_data |>
  filter(quality > 6.5, alcohol < 13.2) |>              #Q7
  arrange(desc(quality)) |>                             #Q8
  select(contains("acid"), alcohol, type, quality) |>   #Q9
  group_by(quality) |>                                  #Q10
  mutate(
    mean_alcohol = mean(alcohol),
    sd_alcohol = sd(alcohol)
  )
head(wine)
```

## Task 3

11. Connect to the database and then look at all of the tables in the database.

```{r}
library(DBI)
library(RSQLite)
library(dplyr)

con <- dbConnect(SQLite(), "data/Lahman.db")

dbListTables(con)
```

12. Use the tbl() function and dplyr to return all of the data from the Teams table for the year 2015.

```{r}
tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()
```

13. Repeat the above by using the sql() function within tbl() (here you have to write actual SQL!).

```{r}
tbl(con, sql("SELECT * FROM Teams WHERE yearID = 2015")) |>
  collect()
```

14. Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the HallOfFame table, the inducted variable is important here.

```{r}
hof <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category) |>
  collect()
head(hof)
```

15. Combine the table from the previous question with the People table in order to have the nameFirst and nameLast names added to what was returned above.

```{r}
people <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast) |>
  collect()

hof_names <- hof |>
  left_join(people, by = "playerID")
head(hof_names)
```

16. Return only the playerID (manager ID in this case), G, W, and L columns from the Managers table. Use chaining to then determine the overall win/loss records for each of the hall of fame managers.

```{r}
managers <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(win_loss_pct = Total_W / G_managed) |>
  arrange(desc(win_loss_pct))
head(managers)
```

17. Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their category in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)

```{r}
hof_managers <- hof_names |>
  inner_join(managers, by = "playerID")

hof_managers
```
