---
title: "Tomas_D"
format: html
editor: visual
---

### Homework 4

#### Task 1

In long format, each measurement gets their own row with each distinct group, variable, or timestamp being its own column

In wide format, each subject gets its own row with repeated measurements spread out across different columns.

For analysis, long format is preferred, it makes data manipulation cleaner and easier with deployed packages.

```{r}
library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
                 1.55), Species)


```

#### Task 2

```{r}
#Reading data and creating column names
data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = c("Id","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type_of_Glass")) |> 
  
    mutate(
      Type_of_Glass = factor(
        Type_of_Glass,
        levels = 1:7,
        labels = c(
          "building_windows_float_processed",
          "building_windows_non_float_processed",
          "vehicle_windows_float_processed",
          "vehicle_windows_non_float_processed",
          "containers",
          "tableware",
          "headlamps"
        )
      )
    ) |>
  
  filter(Fe < 0.2 & Type_of_Glass %in% c("tableware", "headlamps"))
data # printing tibble
```

```{r}
yeast_data <- read.table(
  "https://www4.stat.ncsu.edu/online/datasets/yeast.data", 
  col.names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
) |>
  select(-seq_name, -nuc) |> # remove seq and nuc columns
  group_by(class) |> #add in mutate
  mutate(
    across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(mean = mean, median = median) # add functions
    )
  )

yeast_data


```

```{r}
getwd()
library(readxl) #loading packages

white_wine_s1 <- read_excel("../HW3/data/white-wine.xlsx", sheet =1)#reading data from folder

#white_wine_s1 #viewing tibble

wine_colnames <- read_excel("../HW3/data/white-wine.xlsx",sheet = 2) #Read in second sheet

colnames(white_wine_s1) <- wine_colnames[[1]] #Assign colnames as a vector



#Adding "white" columns

white_wine_s1$type <- "white"

#white_wine_s1

library(readr)
library(dplyr)

#Reading in red wine file
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)

#Assigning column names and new "red" column
colnames(red_wine) <- wine_colnames[[1]]
red_wine$type <- "red"

# print tibble
#red_wine

#combine datasets

wine_data <- bind_rows(white_wine_s1, red_wine)
#wine_data

wine_data_processed <- wine_data |>
  # high quality (> 6.5) and reasonable alcohol (< 13.2)
  filter(quality > 6.5 & alcohol < 13.2) |>
  
  # sort from highest quality to lowest
  arrange(desc(quality)) |>
  
  # select variables containing "acid", plus alcohol, type, and quality
  select(contains("acid"), alcohol, type, quality) |>
  
  # mean and standard deviation of alcohol grouped by quality
  group_by(quality) |>
  mutate(
    alcohol_mean = mean(alcohol),
    alcohol_sd = sd(alcohol)
  ) |>
  ungroup()

# final tibble
wine_data_processed
```

#### Task 3

```{r}
library(DBI)
library(RSQLite)
library(dplyr)

con <- dbConnect(RSQLite::SQLite(), "data/Lahman.db")
dbListTables(con)

#Teams table filter = 2015
teams_2015 <- tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()

#using the sql() 
teams_2015_sql <- tbl(
  con, 
  sql("SELECT * FROM Teams WHERE yearID = 2015")
) |>
  collect()

#playerID, yearID, and category for HoF players
hof_inducted <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

#Combine HoF data with People table, include nameFirst and nameLast
hof_names <- hof_inducted |>
  inner_join(
    tbl(con, "People") |> select(playerID, nameFirst, nameLast), 
    by = "playerID"
  )

#win/loss records for managers and compute win percentage
managers_summary <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  ) |>
  collect() |>
  mutate(win_pct = Total_W / G_managed) |>
  arrange(desc(win_pct))

#Combine manager records with HoF
hof_managers <- inner_join(
  hof_names |> collect(),
  managers_summary,
  by = "playerID"
)

# View final result
hof_managers

dbDisconnect(con)

```

####  

\
\
