---
title: "Homework 2 - Divya Shah"
editor: visual
---

# Task 1: Basic Vector Practice

For this task, I create vectors for the pre-treatment and post-treatment blood pressure measurements. I then calculate the change in blood pressure for each subject and summarize the results.

## 1. Create the vectors

First, I create two vectors containing the pre-treatment and post-treatment blood pressure measurements for the 20 subjects.

```{r}
# Pre-treatment blood pressure measurements
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127,
            141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# Post-treatment blood pressure measurements
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117,
             101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

pre_bp
post_bp
```

## 2. Assign subject names to the vectors

I create subject labels using `paste()` and assign the same names to both blood pressure vectors.

```{r}
# Create subject names
subject_names <- paste("Subject", 1:20, sep = "_")

# Assign names to each vector
names(pre_bp) <- subject_names
names(post_bp) <- subject_names

pre_bp
post_bp
```

## 3. Calculate the change in blood pressure

The change in blood pressure is calculated by subtracting the post-treatment measurement from the pre-treatment measurement. Therefore, a positive value indicates that the subject's blood pressure decreased after treatment.

```{r}
# Calculate change in blood pressure
diff_bp <- pre_bp - post_bp

diff_bp
```

## 4. Calculate the average decrease across all patients

I use `mean()` to calculate the average blood pressure change across all 20 subjects. Since the differences were calculated as pre-treatment minus post-treatment, a positive average represents an overall decrease in blood pressure.

```{r}
# Calculate average blood pressure change
mean(diff_bp)
```

## 5. Determine which patients experienced a decrease

A positive value in `diff_bp` indicates that a patient's blood pressure decreased after treatment. I use `which()` to return the indices and subject names associated with positive changes.

```{r}
# Find subjects with a positive blood pressure change
which(diff_bp > 0)
```

## 6. Return only the positive blood pressure changes

I subset the difference vector so that only subjects with a positive blood pressure change are returned.

```{r}
# Return only positive blood pressure changes
positive_diff <- diff_bp[diff_bp > 0]

positive_diff
```

## 7. Calculate the average decrease for patients whose blood pressure decreased

Finally, I calculate the average decrease only among the subjects whose blood pressure decreased after treatment.

```{r}
# Calculate average decrease among subjects with positive changes
mean(positive_diff)
```

# Task 2: Basic Data Frame Practice

For this task, I organize the blood pressure measurements from Task 1 into a data frame. I then practice subsetting rows and adding a new logical column.

## 1. Create the data frame

I create a data frame with columns for the patient, pre-treatment blood pressure, post-treatment blood pressure, and the calculated difference in blood pressure.

```{r}
# Create the blood pressure data frame
bp_df <- data.frame(
  patient = subject_names,
  pre_bp = unname(pre_bp),
  post_bp = unname(post_bp),
  diff_bp = unname(diff_bp)
)

bp_df
```

## 2. Return rows with a negative blood pressure difference

A negative value in `diff_bp` means that the post-treatment blood pressure was higher than the pre-treatment blood pressure. I subset the data frame using the `diff_bp` column directly.

```{r}
# Return rows where diff_bp is negative
negative_diff <- bp_df[bp_df$diff_bp < 0, ]

negative_diff
```

## 3. Add a column for post-treatment blood pressure below 120

I create a new logical column that is `TRUE` when the post-treatment blood pressure is less than 120 and `FALSE` otherwise.

```{r}
# Add a logical column indicating whether post_bp is below 120
bp_df$post_under_120 <- bp_df$post_bp < 120

bp_df
```

## 4. Display the final data frame

Finally, I use `knitr::kable()` to display the completed data frame in a cleaner table format.

```{r}
# Display the final data frame without row names
knitr::kable(bp_df, row.names = FALSE)
```

# Task 3: List Practice

For this task, I create a second data frame containing blood pressure measurements from a placebo experiment. I then combine the treatment and placebo data frames into a list and practice accessing elements from the list.

## 1. Create the placebo data frame

First, I create vectors containing the pre-treatment and post-treatment blood pressure measurements for the 10 placebo subjects. I then calculate the change in blood pressure and create a data frame similar to the one from Task 2.

```{r}
# Pre-treatment blood pressure measurements for placebo subjects
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

# Post-treatment blood pressure measurements for placebo subjects
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# Create subject names for the placebo group
placebo_subjects <- paste("Subject", 1:10, sep = "_")

# Calculate the change in blood pressure
placebo_diff <- placebo_pre - placebo_post

# Create the placebo data frame
placebo_df <- data.frame(
  patient = placebo_subjects,
  pre_bp = placebo_pre,
  post_bp = placebo_post,
  diff_bp = placebo_diff
)

# Add the same logical column used in Task 2
placebo_df$post_under_120 <- placebo_df$post_bp < 120

placebo_df
```

## 2. Create a list containing both data frames

I create a list with two named elements. The first element, `treatment`, contains the original treatment data frame, and the second element, `placebo`, contains the placebo data frame.

```{r}
# Combine the treatment and placebo data frames into a list
bp_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)

bp_list
```

## 3. Access the first list element using three different types of syntax

There are several ways to access an element from a list in R. Here, I access the treatment data frame using its numeric position, its name with double brackets, and the `$` operator.

```{r}
# Access by numeric position
bp_list[[1]]

# Access by element name using double brackets
bp_list[["treatment"]]

# Access by element name using $
bp_list$treatment
```

## 4. Access the placebo pre-treatment blood pressure column

Finally, I access the `pre_bp` column from the placebo data frame directly from the list in one line.

```{r}
# Access the placebo pre_bp column
bp_list$placebo$pre_bp
```
# Task 4: Control Flow Practice

For this task, I classify each subject's post-treatment blood pressure as `optimal`, `borderline`, or `high`. I add a new `status` column to each data frame stored in the list and use `for` loops with `if`/`else` logic to assign the appropriate category.

## 1. Create the status columns

First, I create an empty character column called `status` in both the treatment and placebo data frames stored inside `bp_list`.

```{r}
# Create empty status columns in the data frames stored in the list
bp_list$treatment$status <- character(nrow(bp_list$treatment))
bp_list$placebo$status <- character(nrow(bp_list$placebo))
```

## 2. Assign status for the treatment group

I loop through each row of the treatment data frame and classify the post-treatment blood pressure. Values less than or equal to 120 are classified as `optimal`, values greater than 120 and less than or equal to 130 are classified as `borderline`, and values greater than 130 are classified as `high`.

```{r}
# Assign blood pressure status for the treatment group
for (i in 1:nrow(bp_list$treatment)) {
  
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
    
  } else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] <- "borderline"
    
  } else {
    bp_list$treatment$status[i] <- "high"
  }
}

bp_list$treatment
```

## 3. Assign status for the placebo group

I repeat the same process for the placebo data frame stored in the list.

```{r}
# Assign blood pressure status for the placebo group
for (i in 1:nrow(bp_list$placebo)) {
  
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
    
  } else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] <- "borderline"
    
  } else {
    bp_list$placebo$status[i] <- "high"
  }
}

bp_list$placebo
```