---
title: "Homework-2"
author: "Javi Brito"
format: html
editor: visual
---

## Task 1: Basic Vector practice

Suppose we have data from a medical experiment on blood pressure. We have the following pre-treatment values for subjects 1 through 20:

- 130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135

after treatment, the subjects were measured again (subjects 1 through 20 match)

- 114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113

1.  Create two vectors. One vector corresponding to the pre measurements and one to the post measurements.

    ```{r t1-p1}
    pre_tr <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

    post_tr <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
    ```

2.  Assign names to the vector elements using the paste() function. Note that names() can be overwritten by a character vector. To quickly create the names, try running the code. Create the same names for each vector’s elements.

    ```{r t1-p2}
    subject_names <- paste("Subject", 1:20, sep = "_")

    names(pre_tr) <- subject_names
    names(post_tr) <- subject_names
    ```

3.  Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements. Recall that R does math element-wise! Save this calculation as a new object (also a vector).

    ```{r t1-p3}
    change_tr <- pre_tr - post_tr
    change_tr
    ```

4.  Calculate the average decrease in blood pressure across all patients.

    ```{r t1-p4}
    avg_change <- mean(change_tr)
    avg_change
    ```

    The average of all 20 patients the blood pressure dropped by 17 points.

5.  Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

    ```{r t1-p5}
    decreased_tr <- which(change_tr > 0)
    decreased_tr
    ```

    Based on the output, we can see how subjects 9,13, and 17 are not part of the list, indicating that their blood pressure went up after the treatment. Thus, the treatment didn't help them.

6.  Subset the vector of differences to only return those that have a positive change.

    ```{r t1-p6}
    decreased_changes_tr <- change_tr[decreased_tr]
    decreased_changes_tr 
    ```

    In this case we know that subjects 9, 13 and 17 are the patients whose blood pressure went up after the treatment.

7.  Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

    ```{r t1-p7}
    avg_change_decreased <- mean(decreased_changes_tr)
    avg_change_decreased
    ```

    Comparing the two resulted averages, notice how when we included everyone, the average decreased by 17, however this resulted included 3 patients whose blood pressure did not decrease. After we calculated the average only including the 17 patients who blood pressure improved, the average decreases about 20.65 points. This results is more accurate to identify the effectiveness of the treatment.

## Task 2: Basic Data Frame Profile

1.  Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp

    ```{r t2-p1}
    patient <- 1:20

    bp_df <- data.frame(patient = patient,
                        pre_bp  = pre_tr,
                        post_bp = post_tr,
                        diff_bp = change_tr,
                        row.names = NULL)
    bp_df
    ```

2.  Return only rows where the diff_bp column is negative. (Use \[ or learn about the subset() function if you’d like. If you use \[, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.)

    ```{r t2-p2}
    subset(bp_df, diff_bp < 0)
    ```

    Notice how these are the same patients in Task 1, whose blood pressure went up after the treatment.

3.  Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120. Recall you can use \$ to reference a column. If you reference a column that doesn’t exist, and save a vector (of appropriate length in it), that vector becomes a column of your data frame! Similar to the previous question, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.

    ```{r t2-p3}
    bp_df$post_under_120 <- bp_df$post_bp < 120
    bp_df
    ```

    The patients 13, 14, and 17 have post_bp values at or above 120.

4.  Finally, print the data frame out nicely in your final document by modifying the code below appropriately

    ```{r t2-p4}
    knitr::kable(bp_df)
    ```

## Task 3: List Practice

Suppose we now also have data from another experiment where the ‘treatment’ was actually a placebo. We have the following pre-treatment values for subjects 1 through 10 (different set of subjects):

- 138, 135, 147, 117, 152, 134, 114, 121, 131, 130

after treatment, the subjects were measured again (subjects 1 through 10 match)

- 105, 136, 123, 130, 134, 143, 135, 139, 120, 124

1.  Create a new data frame with this data in it that is similar to the data frame from task 2 (including the new column).

    ```{r t3-p1}
    pre_tr_pl <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
    post_tr_pl <-c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

    patient_pl <- 1:10
    diff_pl <- pre_tr_pl - post_tr_pl

    bp_df_pl <- data.frame(patient = patient_pl,
                           pre_bp  = pre_tr_pl,
                           post_bp = post_tr_pl,
                           diff_bp = diff_pl,
                           row.names = NULL)

    bp_df_pl$post_under_120 <- bp_df_pl$post_bp < 120
    bp_df_pl
    ```

2.  Now create and store a list with two elements:

    - 1st element named treatment and contains the first data frame you created

    - 2nd element named placebo and contains the second data frame you created.

    ```{r t3-p2}
    experiment_list <- list(treatment = bp_df, placebo = bp_df_pl)
    experiment_list
    ```

3.  Access the first list element using three different types of syntax.

    ```{r t3-p3}
    experiment_list[[1]]
    experiment_list[["treatment"]]
    experiment_list$treatment
    ```

4.  In one line, access the placebo data frame, pre_bp column.

    ```{r t3-p4}
    experiment_list$placebo$pre_bp
    ```

## Task 4: Control Flow Practice

1.  Suppose we want to characterize the post-treatment (or placebo) blood pressure measurement as optimal (≤ 120), borderline (120 \< bp ≤ 130), and high (\> 130). First, create a new column in each data frame from above called status. You can do this via

    `your_df$status <- character(20) #or 10 depending on number of observations`

    Note: You want to add this additional column to the data frames that are stored in your list (R doesn’t do referencing to the original object).

    ```{r t4-p1}
    experiment_list$treatment$status <- character(nrow(experiment_list$treatment))
    experiment_list$placebo$status <- character(nrow(experiment_list$placebo))
    ```

2.  For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.

    ```{r t4-p2}
    for (i in 1:nrow(experiment_list$treatment)) {
      bp_value <- experiment_list$treatment$post_bp[i]
      if (bp_value <= 120) {
        experiment_list$treatment$status[i] <- "optimal"
      } else if (bp_value > 120 & bp_value <= 130) {
        experiment_list$treatment$status[i] <- "borderline"
      } else {
        experiment_list$treatment$status[i] <- "high"
      }
    }

    experiment_list$treatment
    ```

3.  Repeat for the placebo data frame (within the list)

    ```{r t4-p3}
    for (i in 1:nrow(experiment_list$placebo)) {
      bp_value <- experiment_list$placebo$post_bp[i]
      if (bp_value <= 120) {
        experiment_list$placebo$status[i] <- "optimal"
      } else if (bp_value > 120 & bp_value <= 130) {
        experiment_list$placebo$status[i] <- "borderline"
      } else {
        experiment_list$placebo$status[i] <- "high"
      }
    }

    experiment_list$placebo
    ```
