---
title: "Homework 2 - TR"
editor: visual
---

## Task 1 Basic Vector Practice

### Creating pre and post treatment vectors and assigning subject names

```{r}
# Creating Initial Vectors

pretrt <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

posttrt <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

# Vector and patient IDs
subject_ids <- paste("Subject", 1:20, sep = "_")

names(pretrt) <- subject_ids
names(posttrt) <- subject_ids
```

### Calculating the difference between pre & post treatment

```{r}

# initial difference calc
prepostdiff <- c(pretrt - posttrt)

# avg difference
avgdiff <- mean(prepostdiff)

# Subjects where bp decreased
DecreaseBP <- which(prepostdiff > 0)

avgDecrease <- mean(prepostdiff[DecreaseBP]) 

```

### Creating Data Frame

```{r}
#Create data frame no row names
df <- data.frame(patient = subject_ids, pre_bp = pretrt, post_bp = posttrt, diff_bp = prepostdiff, row.names = NULL)

negative_diff <- df[df$diff_bp < 0, ]

df$Post_Under120 <- df$post_bp < 120

knitr::kable(df)
```

## Task 3 List Practice

### Creating lists for placebo dataset, replicating dataframe

```{r}

#placebo vectors
pretrt_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

posttrt_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo_diff <- pretrt_placebo - posttrt_placebo

#placebo data frame creation
placebo_df <- data.frame(patient = subject_ids, pre_bp = pretrt_placebo, post_bp = posttrt_placebo, diff_bp = placebo_diff, row.names = NULL )

placebo_df$Post_Under120 <- placebo_df$post_bp < 120
#render
knitr::kable(placebo_df)

# Creating lists
list_bp <- list(treatment = df, placebo = placebo_df)

list_bp$treatment
list_bp[[1]]
list_bp[["treatment"]]

list_bp$placebo$pre_bp
```

## Task 4 Control Flow Practice

### Loop logic for creating new columns for categories for non-placebo data

```{r}
# Creating new character columns for categories 
list_bp$treatment$status <- character(20)

list_bp$placebo$status <- character(10)

# For loop logic for categorizing bp levels
for(i in 1:nrow(list_bp$treatment)){
  
  if (list_bp$treatment$post_bp[i] <= 120) {list_bp$treatment$status[i] <- "optimal"}
  
  else if (list_bp$treatment$post_bp[i] > 120 & list_bp$treatment$post_bp[i] <=130) {list_bp$treatment$status[i] <- "borderline"}
  
  else {list_bp$treatment$status[i] <- "high"}
}
# render
knitr::kable(list_bp$treatment)


```

### Loop logic for the placebo dataset, creating buckets for post-placebo bp levels.

```{r}
# For loop logic for categorizing bp levels
for(i in 1:nrow(list_bp$placebo)){
  
  if (list_bp$placebo$post_bp[i] <= 120) {list_bp$placebo$status[i] <- "optimal"}
  
  else if (list_bp$placebo$post_bp[i] > 120 & list_bp$placebo$post_bp[i] <=130) {list_bp$placebo$status[i] <- "borderline"}
  
  else {list_bp$placebo$status[i] <- "high"}
}

#render
placebo_df_status <- list_bp$placebo
knitr::kable(list_bp$placebo)


```
