---
title: "Homework 2 - Maab Aldulimy"
editor: visual
---

## Task 1: Basic Vector Practice

In this section, I create vectors for pre-treatment and post-treatment, then calculate difference between the respective vectors to determine average difference.

```{r}
#creating needed lists of pre and post treatment blood pressure measurements:
pre_treatment<- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_treatment<- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

# Assigning names to the vector elements using the paste() function.

subject_names <-paste("Subject", 1:20, sep = "_")

#using names() function 
names(pre_treatment) <- subject_names
names(post_treatment) <- subject_names

#Calculating the change in blood pressure across all patients.
bp_change<- pre_treatment- post_treatment
bp_change

# Calculating the average decrease in blood pressure across all patients.
avg_dec <- mean(bp_change)
avg_dec

#finding which patients experienced a decrease in blood pressure after treatment (a positive change)
find_pos_change<-which(bp_change > 0)
find_pos_change
# selecting only positive blood pressure changes
pos_change <- bp_change[bp_change >0]
pos_change

#average decrease in blood pressure for positive change valus 
mean(pos_change)
```

## Task 2: Basic Data Frame Practice

In this section I build a data frame from the previous vectors, filter for negative differences in blood pressure, add a normal range column which evaluates if blood pressure is less than 120, and print this data frame using the kable() function.

```{r}
#creating a data frame with columns patient,pre_bp,post_bp,diff_bp
my_data <- data.frame(patient=subject_names,
                      pre_bp=pre_treatment,
                      post_bp=post_treatment,
                      diff_bp=bp_change,
                      row.names = NULL)
my_data

# Returning only rows where the diff_bp column is negative

neg_diff_bp <- my_data[my_data$diff_bp < 0,]  # form my_data[row,columns]
neg_diff_bp

#Adding a new column named normal_range to the data frame that is TRUE if the post_bp is less than 120
my_data$ normal_range <- my_data$post_bp < 120

knitr::kable(my_data)

```

## Task 3: List Practice

In this section I build the placebo data frame (same structure as Task 2), then combine both data frames into a named list, and demonstrate a few ways to access elements from that list.

```{r}
#creating placebo vectors 
Pre_placebo<- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
Post_placebo<- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# Assigning names to the placebo vector elements using the paste() function.

p_subject_names <-paste("Subject", 1:10, sep = "_")

#using names() function 
names(Pre_placebo) <- p_subject_names
names(Post_placebo) <- p_subject_names

#Calculating the change in blood pressure across all patients.
p_bp_change<- Pre_placebo- Post_placebo
p_bp_change

#creating a placebo data frame with columns patient,pre_bp,post_bp,diff_bp
p_my_data <- data.frame(patient=p_subject_names,
                      pre_bp=Pre_placebo,
                      post_bp=Post_placebo,
                      diff_bp=p_bp_change,
                      normal_range=Post_placebo < 120,
                      row.names = NULL)
p_my_data

#creating a list with two elements : treatment and placebo 
bp_list <- list(treatment = my_data, placebo = p_my_data)

#Three different ways to access the first element treatment form the bp_list:
bp_list$treatment
bp_list[["treatment"]] # using double brackets to get same object type data frame each time
bp_list[[1]]

# In one line, accessing the placebo data frame, pre_bp column.
bp_list[["placebo"]]$pre_bp
```

## Task 4: Control Flow Practice

In this section I add a status column to each data frame inside the list, then use a for loop with if/else logic to classify each patient as optimal, borderline, or high based on their post treatment/placebo blood pressure values.

```{r}
#creating a status column in treatment and placebo:
bp_list$treatment$status<- character(20)
bp_list$placebo$status<- character(10)

#creating loop for treatment data frame with conditions optimal (≤ 120), borderline (120 < bp ≤ 130), and high (> 130).
for (i in 1:nrow(bp_list$treatment)){
  
  if(bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i]<- "optimal" 
    
  }else if (bp_list$treatment$post_bp[i] > 120 & bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i]<- "borderline" 
    
  }else {
    bp_list$treatment$status[i]<- "high" 
  }
}

bp_list$treatment

#creating the same loop with conditions optimal (≤ 120), borderline (120 < bp ≤ 130), and high (> 130) for placebo 

for ( i in 1:nrow(bp_list$placebo)){
  
  if (bp_list$placebo$post_bp[i] <= 120){
    bp_list$placebo$status[i]<- "optimal"
    
  }else if (bp_list$placebo$post_bp[i] > 120 & bp_list$placebo$post_bp[i]<= 130){
    bp_list$placebo$status[i]<-"borderline"
    
  }else {
    bp_list$placebo$status[i]<-"high"
  }
}
bp_list$placebo
```
