---
title: "Student E"
author: "Madhu"
editor: visual
---

## Task 1

Given are pre treatment values \[130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135\] and post treatment values \[114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113\] for patients 1 to 20

#### 1. Create 2 vectors

```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

#### 2. Assign elements using paste()

```{r}
name_list <- paste("Subject", 1:20, sep ="_")

names(pre) <- name_list
names(post) <- name_list

```

#### 3. Calculate $\Delta$ BP

```{r}

delta <- pre-post
delta
```

#### 4. Calculate avg $\Delta$ BP

```{r}
mean(delta)
```

#### 5. Determine which patients experienced a decrease in BP (+ $\Delta$). Use which() and return associated indices and names

```{r}
decreased_bp <- which (delta > 0)
decreased_bp
```

#### 6. Vector with only positive $\Delta$ BP

```{r}
positive <- delta[decreased_bp]
positive
```

#### 7. Calculate avg positive $\Delta$ BP

```{r}
mean(positive)
```

## Task 2

#### 1. Create a data frame object with four columns: patient, pre_bp, post_bp, and diff_bp

```{r}
bp_df <- data.frame(patient = name_list, 
                              pre_bp = unname(pre), 
                              post_bp = unname(post), 
                              diff_bp = unname(delta))
```

#### 2. Return $-\Delta$ BP rows

```{r}
subset(bp_df, diff_bp < 0)
```

#### 3. Add boolean column indicating if post_bp \< 120

```{r}

bp_df$low_post_bp <- bp_df$post_bp < 120

```

#### 4. Print the data frame

```{r}
knitr::kable(bp_df)
```

## Task 3

New "treatment" except its a placebo for 10 patients

#### 1. Create new df similar to task 2

```{r}
pre_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo_bp_df <- data.frame(patient = paste("Subject", 1:10, sep ="_"),
                            pre_bp = pre_placebo,
                            post_bp = post_placebo,
                            diff_bp = pre_placebo - post_placebo,
                            low_post_bp = post_placebo < 120
                            )

```

#### 2. Create and store a list with two elements: treatment, placebo

```{r}
bp_list <- list(treatment = bp_df, placebo = placebo_bp_df)
bp_list
```

#### 3. Access the first list element using three different types of syntax

```{r}
bp_list[[1]]
bp_list[["treatment"]]
bp_list$treatment
```

#### 4. Access the placebo data frame, pre_bp column

```{r}
bp_list$placebo$pre_bp
```

## Task 4

#### Categorize post bp (optimal ≤ 120, borderline 120 \< bp ≤ 130, high \> 130) 

#### 1. Create new column status

```{r}
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```

#### 2. Categorize non-placebo with for loops and if/then/else statements

```{r}
for (i in 1:20) {
  
  x <- bp_list$treatment$post_bp[i]
  
  if (x <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  } 
  
  else if (x > 130) {
    bp_list$treatment$status[i] <- "high"
  } 
  
  else {
    bp_list$treatment$status[i] <- "borderline"
  }
  
}
```

#### 3. Repeat for placebo data

```{r}

for (i in 1:10) {
  
  x <- bp_list$placebo$post_bp[i]
  
  if (x <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  } 
  
  else if (x > 130) {
    bp_list$placebo$status[i] <- "high"
  } 
  
  else {
    bp_list$placebo$status[i] <- "borderline"
  }
  
}
```
