---
title: "Divya_A"
editor: visual
---

## Task 1: Conceptual Questions

1.  **If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?**\
    The relative path would be `../MyData.csv`.

2.  **What are the major benefits of using R projects?**\
    R projects help keep files organized, make it easier to use relative file paths, and keep project-specific files and settings together.

3.  **What is Git and what is GitHub?**\
    Git is a version control system that tracks changes to files. GitHub is an online platform used to store Git repositories and collaborate with others.

4.  **What are the two main differences between a tibble and a data.frame?**\
    Tibbles print a smaller preview of the data instead of printing everything, and they are stricter when accessing columns and do not partially match column names.

## Task 2: Reading Delimited Data

### Glass Data

The glass data is comma-delimited and does not include column names. I read the data directly from the URL and added the appropriate variable names.

```{r}
library(tidyverse)

glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "Id", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "Type"
  )
)

glass
```

### Yeast Data

The yeast data is separated by whitespace and does not include column names. I read the data directly from the URL and added the appropriate variable names.

```{r}
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  )
)

yeast
```

## Task 3: Combining Excel and Delimited Data

### White Wine Data

First, I read the white wine data from the first sheet of the Excel file using the `readxl` package.

```{r}
#| message: false
#| warning: false

library(readxl)
library(tidyverse)

white_wine <- read_excel(
  "data/white-wine.xlsx",
  sheet = 1
)

white_wine
```

Next, I read the alternative variable names from the second sheet and use them to replace the current column names.

```{r}
white_names <- read_excel(
  "data/white-wine.xlsx",
  sheet = 2
)

colnames(white_wine) <- white_names[[1]]

white_wine
```

I then add a column indicating that every observation in this data set is a white wine.

```{r}
white_wine <- white_wine |>
  mutate(type = "white")

white_wine
```

### Red Wine Data

The red wine data is semicolon-delimited. I read it directly from the URL using `readr`, replace the column names with the same variable names used for the white wine data, and add a column identifying the observations as red wine.

```{r}
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";",
  col_types = cols(.default = col_double())
)

colnames(red_wine) <- white_names[[1]]

red_wine <- red_wine |>
  mutate(type = "red")

red_wine
```

### Combined Wine Data

Finally, I combine the white and red wine data into one tibble using `bind_rows()`.

```{r}
wine_data <- bind_rows(
  white_wine,
  red_wine
)

wine_data
```

## Task 4: Writing Functions

### Interquartile Range Function

First, I create a function called `getIQR` that calculates the interquartile range by subtracting the 25th percentile from the 75th percentile. The returned value is unnamed, and additional arguments can be passed to `quantile()`.

```{r}
getIQR <- function(x, ...) {
  result <- quantile(x, 0.75, ...) - quantile(x, 0.25, ...)
  unname(result)
}
```

I apply the function to the `fixed_acidity` and `citric_acid` columns of the combined wine data.

```{r}
getIQR(wine_data$fixed_acidity)

getIQR(wine_data$citric_acid)
```

### Coefficient of Variation Function

Next, I create a function called `getCoefVar` that calculates the sample coefficient of variation as the sample standard deviation divided by the absolute value of the sample mean.

```{r}
getCoefVar <- function(x, ...) {
  sd(x, ...) / abs(mean(x, ...))
}
```

I apply this function to the same two variables.

```{r}
getCoefVar(wine_data$fixed_acidity)

getCoefVar(wine_data$citric_acid)
```

### Summary Function

Next, I create a wrapper function called `getSummaries`. This function calls the two functions created above and returns both the IQR and coefficient of variation with appropriate names.

```{r}
getSummaries <- function(x, ...) {
  c(
    IQR = getIQR(x, ...),
    CoefVar = getCoefVar(x, ...)
  )
}
```

I apply the summary function to the `fixed_acidity` and `citric_acid` variables.

```{r}
getSummaries(wine_data$fixed_acidity)

getSummaries(wine_data$citric_acid)
```

### Infix Function

Finally, I create an infix function that can be used in place of `cbind()`.

```{r}
`%cbind%` <- function(x, y) {
  cbind(x, y)
}
```

I create the provided data frames and vector and test the infix function first with the two data frames and then with all three objects.

```{r}
set.seed(1)

df1 <- data.frame(
  unif = runif(10),
  norm = rnorm(10)
)

df2 <- data.frame(
  gamma = rgamma(10, shape = 1),
  beta = rbeta(10, shape1 = 5, shape2 = 1)
)

vec1 <- rnorm(10, mean = 10)

# Combine the two data frames
df1 %cbind% df2

# Combine both data frames and the vector
(df1 %cbind% df2) %cbind% vec1
```
