---
title: "HW 3"
format: html
editor: visual
---

## Task 1

1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

    - `../MyData.csv`, since `..` indicates the parent folder

2.  What are the major benefits of using R projects?

    - R projects is an easy way to divide and organize your work. Each project contains a working directory, its own workspace, history, as well as folder structure & source documents.

3.  What is Git and what is GitHub?

    - Git is a version control software, used so that we can periodically save snapshots of our files/work, whereas GitHub is the online platform that stores our git repositories.
    - GitHub is also key for collaboration with others, so that multiple people can share snapshots of repos and work on them at the same time and push it to the platform.

4.  What are the two main differences between a tibble and a data.frame?

    - Tibble is similar to a data.frame, but it is a modern re-imagining that's more compatible for the wonderful tidyverse library.
    - They have a more compact print, and complain more upfront leading to cleaner code.
    - Subsetting one column using \[ preserves the tibble, but df forces it down to a vector.

## Task 2

1.  Read glass.data into R from URL using `tidyverse`

    ```{r}
    library(tidyverse)

    glass_data <- read_csv(
      "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
      col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")
    )
    glass_data


    ```

2.  Read yeast.data into R from URL using `tidyverse`

    ```{r}
    yeast_data <- read_table(
      "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
      col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
    )

    yeast_data
    ```

glass.data's delimiter were commas and yeast.data's were whitespaces.

## Task 3

3.  Download wine.data from URL, import first sheet using `readxl`,

    ```{r}
    library(readxl)
    library(tidyverse)

    white_wine <- read_excel("data/white-wine.xlsx", sheet = 1)

    white_wine
    ```

4.  Read in data from second sheet, overwrite the current column names using this data.

    ```{r}
    new_names <- read_excel("data/white-wine.xlsx", sheet = 2)

    colnames(white_wine) <- new_names[[1]]
    white_wine
    ```

5.  Add a column stating the type of wine (all are white)

    ```{r}
    white_wine$type<-"white"
    ```

6.  Import red wine data from URL using `readr` (semi colon delimiter), replace column names, add a column stating the type (red)

    ```{r}
    red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim=";")

    colnames(red_wine) <- new_names[[1]]

    red_wine$type <- "red"
    ```

7.  Combine the 2 tibbles

    ```{r}
    wine <- bind_rows(white_wine, red_wine)
    wine
    ```

## Task 4

8.  Write a basic function (call it `getIQR`) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the quantile function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the quantile function.

    ```{r}
    getIQR <- function(x, ...) {
      q <- quantile(x, probs = c(0.25, 0.75), ...)
      unname(q[2] - q[1])
    }
    ```

9.  Take the object from Q7 and apply the getIRQ() function to the `fixed_acidity` and `citric_acid` columns (two function calls).

    ```{r}
    getIQR(wine$fixed_acidity)
    getIQR(wine$citric_acid)
    ```

10. Write another function (call it `getCoefVar`) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean.

    ```{r}
    getCoefVar <- function(x, ...) {
      sd(x, ...) / abs(mean(x, ...))
    }
    ```

11. Repeat Q9 using the getCoefVar() function.

    ```{r}
    getCoefVar(wine$fixed_acidity)
    getCoefVar(wine$citric_acid)
    ```

12. Create a function (call it `getSummaries`) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a wrapper function.)

    ```{r}
    getSummaries <- function(x, ...) {
      c(IQR = getIQR(x, ...),
        CoefVar = getCoefVar(x, ...))
    }
    ```

13. Repeat question 2 with the `getSummaries()` function.

    ```{r}
    getSummaries(wine$fixed_acidity)
    getSummaries(wine$citric_acid)
    ```

14. Write your own infix function that can be used in replace of the cbind function (feel free to call the cbind function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

    ```{r}
    df1 <- data.frame(unif = runif(10), norm = rnorm(10))
    df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
    vec1 <- rnorm(10, mean = 10)
    df1
    df2
    vec1
    ```

    New infix defined below

    ```{r}
    `%cb%` <- function(x, y) {
      cbind(x, y)
    }
    ```

    Check using given dfs

    ```{r}
    df1 %cb% df2

    df1 %cb% df2 %cb% vec1
    ```
