---
title: "Tomas_D"
format: html
editor: visual
---

## Homework 3

#### Task 1

1\. If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

- ../MyData.csv

2\. What are the major benefits of using R projects?

- The main benefits of R projects include relative filepaths and working directories to quickly access relevant scripts and data. Also, there are isolated workspaces and easy version control implementation.

3\. What is Git and what is GitHub?

- Git is a version control software used in project development. Github is an online repository for the same projects, allowing collaboration, sharing, and an easier GUI for navigation.

4\. What are the two main differences between a tibble and a data.frame?

- When printing a data.frame, all rows and columns are shown, vs a tibble only the first 10 rows are shown, in a more compact view. Tibbles will only return tibbles when sub-setting a column, dfs will return a vector.

#### Task 2

```{r}
library(tidyverse) #loading packages


#Reading data and creating column names
data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = c("Id","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type_of_Glass"))

data # printing tibble
```

```{r}
yeast_data <- read.table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col.names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))

#yeast_data
```

#### Task 3

```{r}
getwd() # finding my place
library(readxl) #loading packages

white_wine_s1 <- read_excel("data/white-wine.xlsx", sheet =1)#reading data from folder

white_wine_s1 #viewing tibble

wine_colnames <- read_excel("data/white-wine.xlsx",sheet = 2) #Read in second sheet

colnames(white_wine_s1) <- wine_colnames[[1]] #Assign colnames as a vector



#Adding "white" columns

white_wine_s1$type <- "white"

white_wine_s1

```

##### Task 3 Part 2

```{r}
#Loading Packages
library(readr)
library(dplyr)

#Reading in red wine file
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)

#Assigning column names and new "red" column
colnames(red_wine) <- wine_colnames[[1]]
red_wine$type <- "red"

# print tibble
red_wine

#combine datasets

wine_data <- bind_rows(white_wine_s1, red_wine)
wine_data
```

#### Task 4

```{r}
#Define function and arguments
getIQR <- function(x,...){
  
  #Compute quantiles
  quantiles <- quantile(x, probs = c(0.25,0.75),...)
  
  #calc iqr
  iqr <- quantiles[2] - quantiles[1]
  
  return(unname(iqr))
  
}

fixed_acidity_iqr <- getIQR(wine_data$fixed_acidity)
citric_acid_iqr <- getIQR(wine_data$citric_acid)

fixed_acidity_iqr
citric_acid_iqr
```

```{r}
#Define function
getCoefVar <- function(x,...){
  
  #sd function
  sd <- sd(x,...)
  
  #mean function
  mean <- mean(x,...)
  
  #cv calc
  cv <- sd/ abs(mean)
  
  return(cv)
}

#Calc CV values
fixed_acidity_cv <- getCoefVar(wine_data$fixed_acidity)
citric_acid_cv <- getCoefVar(wine_data$citric_acid)

fixed_acidity_cv
citric_acid_cv
```

```{r}
#Define function
getSummaries <- function(x,...){
  iqr_calc <- getIQR(x,...)
  iqr_cv <- getCoefVar(x,...)
  
  out <- c(IQR = iqr_calc, CV = iqr_cv)
  
  return(out)
}

#Calc summaries
fixed_acidity_summary <- getSummaries(wine_data$fixed_acidity)
citric_acid_summary <- getSummaries(wine_data$citric_acid)

fixed_acidity_summary
citric_acid_summary
```

```{r}
#Infix operator

`%cb%` <- function(a, b) {
  cbind(a, b)
}

#Set random seeds 
set.seed(1)

#create set data frames
df1  <- data.frame(unif = runif(10), norm = rnorm(10))
df2  <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

#testing operators
test1 <- df1 %cb% df2

print(test1)

test2 <- df1 %cb% df2 %cb% vec1

print(test2)
```
