---
title: "Homework 5 - Divya"
author: "Divya Shah"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(httr)
library(jsonlite)

api_key <- Sys.getenv("NEWS_API_KEY")

if (api_key == "") {
  stop("NEWS_API_KEY has not been set.")
}
```


## Task 1 - Querying an API and a Tidy-Style Function

### 1. Query the News API

For my first query, I searched for recent news about artificial intelligence.

```{r}
news_request <- GET(
  "https://newsapi.org/v2/everything",
  query = list(
    q = "artificial intelligence",
    from = "2026-09-01",
    language = "en",
    pageSize = 50,
    apiKey = api_key
  )
)

stop_for_status(news_request)
```

The request is stored in the `news_request` object.


### 2. Parse the API Response

I parse the returned data and extract the article information into a tibble. The first column, `source`, is kept as a list column.

```{r}
parsed_news <- content(
  news_request,
  as = "parsed",
  type = "application/json"
)

names(parsed_news)
```

```{r}
get_value <- function(x) {
  if (is.null(x)) {
    NA_character_
  } else {
    x
  }
}

article_data <- map_dfr(
  parsed_news$articles,
  function(article) {
    tibble(
      source = list(article$source),
      author = get_value(article$author),
      title = get_value(article$title),
      description = get_value(article$description),
      url = get_value(article$url),
      urlToImage = get_value(article$urlToImage),
      publishedAt = get_value(article$publishedAt),
      content = get_value(article$content)
    )
  }
)

glimpse(article_data)
```

I display the first few articles below.

```{r}
article_data |>
  select(source, author, title, publishedAt) |>
  slice_head(n = 5)
```


### 3. Create a Function to Query the API

The following function allows a user to provide a subject, starting date, and API key. It returns the article information as a tibble.

```{r}
query_news <- function(subject, from_date, api_key) {
  
  response <- GET(
    "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = from_date,
      language = "en",
      pageSize = 50,
      apiKey = api_key
    )
  )
  
  stop_for_status(response)
  
  parsed <- content(
    response,
    as = "parsed",
    type = "application/json"
  )
  
  articles <- map_dfr(
    parsed$articles,
    function(article) {
      tibble(
        source = list(article$source),
        author = get_value(article$author),
        title = get_value(article$title),
        description = get_value(article$description),
        url = get_value(article$url),
        urlToImage = get_value(article$urlToImage),
        publishedAt = get_value(article$publishedAt),
        content = get_value(article$content)
      )
    }
  )
  
  return(articles)
}
```

I use the function once to search for artificial intelligence news.

```{r}
ai_news <- query_news(
  subject = "artificial intelligence",
  from_date = "2026-09-01",
  api_key = api_key
)

ai_news |>
  select(title, publishedAt) |>
  slice_head(n = 5)
```

I use the function a second time to search for healthcare news.

```{r}
health_news <- query_news(
  subject = "healthcare",
  from_date = "2026-09-01",
  api_key = api_key
)

health_news |>
  select(title, publishedAt) |>
  slice_head(n = 5)
```


## Task 2 - Summarizing Roller Coaster Data

### Read and Prepare the Data

I read in the roller coaster data and convert appropriate categorical variables to factors.

```{r}
coasters <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv",
  na = c("", "NA"),
  show_col_types = FALSE
)

names(coasters) <- make.names(
  names(coasters),
  unique = TRUE
)

coasters <- coasters |>
  mutate(
    State = as.factor(State),
    Census.Region = as.factor(Census.Region),
    AgeGroup = factor(
      AgeGroup,
      levels = c("1:Older", "2:Recent", "3:Newest"),
      labels = c("Older", "Recent", "Newest")
    ),
    Scale = as.factor(Scale),
    Type = as.factor(Type),
    Design = as.factor(Design),
    Inversions. = as.factor(Inversions.),
    Make = as.factor(Make)
  )
```


### 4. Examine How the Data Is Stored

I inspect the dimensions, variable names, and structure of the data.

```{r}
dim(coasters)
```

```{r}
names(coasters)
```

```{r}
glimpse(coasters)
```

The data contains both numeric and categorical variables. Variables representing categories, such as coaster type, scale, design, and inversion status, have been converted to factors.


### 5. Missing Values

I count the number of missing values in every variable.

```{r}
missing_values <- coasters |>
  summarise(
    across(
      everything(),
      ~sum(is.na(.))
    )
  )

missing_values
```

This allows us to see which variables contain missing observations and how many values are missing from each variable.


## Categorical Variables

### 6. Contingency Tables

#### One-Way Table

I create a one-way contingency table for roller coaster type.

```{r}
one_way <- table(coasters$Type)

one_way
```

There are `r one_way["Steel"]` steel roller coasters in the data.


#### Two-Way Table

I create a two-way contingency table comparing coaster type and scale.

```{r}
two_way <- table(
  coasters$Type,
  coasters$Scale
)

two_way
```

The table shows that `r two_way["Steel", "Extreme"]` roller coasters are both steel and classified as extreme.


#### Three-Way Table

I create a three-way contingency table using coaster type, scale, and age group.

```{r}
three_way <- table(
  coasters$Type,
  coasters$Scale,
  coasters$AgeGroup
)

three_way
```

There are `r three_way["Steel", "Extreme", "Newest"]` roller coasters that are steel, classified as extreme, and in the newest age group.


### 7. Conditional Two-Way Tables

I condition on only the newest roller coasters and create a two-way table of type and scale.

#### Method 1: Filter the Data First

```{r}
newest_coasters <- coasters |>
  filter(AgeGroup == "Newest")

conditional_table_1 <- table(
  newest_coasters$Type,
  newest_coasters$Scale
)

conditional_table_1
```

#### Method 2: Subset the Three-Way Table

```{r}
conditional_table_2 <- three_way[, , "Newest"]

conditional_table_2
```

Both methods return the type-by-scale contingency table for only the newest roller coasters.


### 8. Two-Way Table Using dplyr

I recreate a two-way contingency table using `group_by()` and `summarise()`.

```{r}
two_way_dplyr <- coasters |>
  filter(
    !is.na(Type),
    !is.na(Scale)
  ) |>
  group_by(Type, Scale) |>
  summarise(
    Count = n(),
    .groups = "drop"
  )

two_way_dplyr
```

I then use `pivot_wider()` to make the result resemble the output from `table()`.

```{r}
two_way_wide <- two_way_dplyr |>
  pivot_wider(
    names_from = Scale,
    values_from = Count,
    values_fill = 0
  )

two_way_wide
```


### 9. Bar Graphs

#### Stacked Bar Graph

```{r}
coasters |>
  filter(
    !is.na(Type),
    !is.na(Scale)
  ) |>
  ggplot(
    aes(
      x = Type,
      fill = Scale
    )
  ) +
  geom_bar(position = "stack") +
  labs(
    title = "Roller Coaster Scale by Type",
    x = "Coaster Type",
    y = "Number of Roller Coasters",
    fill = "Scale"
  )
```

#### Side-by-Side Bar Graph

```{r}
coasters |>
  filter(
    !is.na(Type),
    !is.na(Scale)
  ) |>
  ggplot(
    aes(
      x = Type,
      fill = Scale
    )
  ) +
  geom_bar(position = "dodge") +
  labs(
    title = "Roller Coaster Scale by Type",
    x = "Coaster Type",
    y = "Number of Roller Coasters",
    fill = "Scale"
  )
```