---
title: "HW 5"
format: html
---

## Task 1: Querying an API and a Tidy-Style Function1

1.  Use `GET()` from `httr` package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r}
library(httr)

api_key <- Sys.getenv("API_KEY")

output <- GET(
  paste0("https://newsapi.org/v2/everything?q=%22black%20hole%22&from=2026-09-01&language=en&pageSize=50&apiKey=", api_key)
)

```

2.  Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
library(jsonlite)
library(tidyverse)

parsed <- fromJSON(rawToChar(output$content))
news_info <- as_tibble(parsed$articles)

str(news_info)
news_info
```

3.  Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
news_query_function <- function(subject, from_date, key) {
  
  output <- GET(
    paste0(
      "https://newsapi.org/v2/everything?q=",
      URLencode(subject),
      "&from=",
      from_date,
      "&language=en&pageSize=50&apiKey=",
      key
      )
    )
  
  parsed <- fromJSON(rawToChar(output$content))
  
  as_tibble(parsed$articles)
}

```

Try it out 1

```{r}

news_query_function(subject = "project 2025", from_date = "2026-08-30", key = api_key)

```

Try it out 2

```{r}
news_query_function(subject = "nutrition", from_date = "2026-08-31", key = api_key)
```

## Task 2: Summarizing Roller Coaster Data

```{r}
library(janitor)
```

```{r}
coasters <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
) |> 
  clean_names() |>
  mutate(
    age_group = factor(age_group),
    census_Region = factor(census_region),
    scale = factor(scale),
    type = factor(type),
    design = factor(design),
    make = factor(make),
    state = factor(state),
    inversions = factor (inversions)
  )

```

State, Make, Design, Type, Scale, Census Region, AgeGroup, Scale, Inversions were converted to factors.

4.  Look at how the data is stored and see if everything makes sense

```{r}
glimpse(coasters)
summary(coasters)
```

Its easy to see the summary stats (spread etc) and what type of data the columns are using `glimpse()` and `summary()`. Specifically the factors are likely easier to use when creating the contingency tables.

5\. Document any missing values in the data

```{r}
colSums(is.na(coasters))
```

`TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, Make have missing values, and the `Drop` column has the most missing values by far.

6.  Create a one way contingency, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this.
    - Interpret a number from each resulting table (that is, pick out a value produced and explain what that value means.)

```{r}
# One-way
table(coasters$type)

```

There are 105 total wood coasters in the dataset.

```{r}
# Two-way
table(coasters$type, coasters$census_region)
```

There are 30 wooden coasters in the Northeast Census region in the data set.

```{r}
# Three-way
table(coasters$type, coasters$census_region, coasters$scale)
```

There are 3 wooden coasters in the Census defined South region classified as family level thrill.

7.  Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods:

- Once, by subsetting the data (say with filter()) and then creating the two-way table

```{r}
coasters |>
  filter(scale == "Thrill") |>
  with(table(type, census_region))
```

- Once, by creating a three-way table and subsetting it

```{r}
three_way_subsect <- table(
  coasters$type,
  coasters$census_region,
  coasters$scale
)

three_way_subsect <- three_way_subsect[, , "Thrill"]
three_way_subsect
```

8.  Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result look more like the output from table().

```{r}
coasters |>
  group_by(type, census_region) |>
  summarize(count = n(), .groups = "drop") |>
  pivot_wider(
    names_from = census_region,
    values_from = count,
    values_fill = 0
  )
```

Looks the same as two way contingency table from 6, except it's a tibble

9.  Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

```{r}
#stacked
ggplot(coasters, aes(x = type, fill = census_region)) +
  geom_bar() +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Roller Coaster Type by Census Region"
  )

#side by side
ggplot(coasters, aes(x = type, fill = census_region)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Number of Roller Coasters by Census Region"
  )

```
