---
title: "HW 5"
author: "Madhu"
format: html
---

## Task 1

1.  Use Get() from `httr` package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r}
library(httr)

api_key <- Sys.getenv("API_KEY")

output <- GET(
  paste0("https://newsapi.org/v2/everything?q=%22black%20hole%22&from=2026-09-01&language=en&pageSize=50&apiKey=", api_key)
)

```

2.  Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
library(jsonlite)
library(tidyverse)

parsed <- fromJSON(rawToChar(output$content))
news_info <- as_tibble(parsed$articles)

news_info
```

3.  Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
news_query_function <- function(subject, from_date, key) {
  
  output <- GET(
    paste0(
      "https://newsapi.org/v2/everything?q=",
      URLencode(subject),
      "&from=",
      from_date,
      "&language=en&pageSize=50&apiKey=",
      key
      )
    )
  
  parsed <- fromJSON(rawToChar(output$content))
  
  as_tibble(parsed$articles)
}

```

Try it out 1

```{r}

news_query_function(subject = "project 2025", from_date = "2026-08-30", key = api_key)

```

Try it out 2

```{r}
news_query_function(subject = "nutrition", from_date = "2026-08-31", key = api_key)
```

## Task 2

```{r}
coasters <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)

coasters <- coasters |>
  mutate(
    AgeGroup = factor(AgeGroup),
    `Census Region` = factor(`Census Region`),
    Scale = factor(Scale),
    Type = factor(Type),
    Design = factor(Design),
    Make = factor(Make)
  )
```

4.  Look at how the data is stored and see if everything makes sense

```{r}
glimpse(coasters)
summary(coasters)
```
Its easy to see the summary stats (spread etc) and what type of data the columns are using `glimpse()` and `summary()`. Specifically the factors are likley easier to use when creating the contingency tables.
5.  Document any missing values in the data

```{r}
colSums(is.na(coasters))
```

TopSpeed, MaxHeight, Drop, Length, Duration, Make have missing values

6.  Create a one way contingency, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this.
    - Interpret a number from each resulting table (that is, pick out a value produced and explain what that value means.)

```{r}
# One-way
table(coasters$Type)

```
There are 105 total wood coasters in the dataset.

```{r}
# Two-way
table(coasters$Type, coasters$`Census Region`)
```
There are 30 wooden coasters in the Northeast Census region in the data set.

```{r}
# Three-way
table(coasters$Type, coasters$`Census Region`, coasters$Scale)
```
There are 3 wooden coasters in the Census defined South region classified as family level thrill.

7.  Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods:

- Once, by subsetting the data (say with filter()) and then creating the two-way table

```{r}
coasters |>
  filter(Scale == "Thrill") |>
  with(table(Type, `Census Region`))
```

- Once, by creating a three-way table and subsetting it

```{r}
three_way_subsect <- table(
  coasters$Type,
  coasters$`Census Region`,
  coasters$Scale
)

three_way_subsect <- three_way_subsect[, , "Thrill"]
three_way_subsect
```

8.  Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result look more like the output from table().

```{r}
coasters |>
  group_by(Type, `Census Region`) |>
  summarize(count = n(), .groups = "drop") |>
  pivot_wider(
    names_from = `Census Region`,
    values_from = count,
    values_fill = 0
  )
```
Looks the same as two way contingency table from 6, except it's a tibble

9.  Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

```{r}
#stacked
ggplot(coasters, aes(x = Type, fill = `Census Region`)) +
  geom_bar() +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Roller Coaster Type by Census Region"
  )

#side by side
ggplot(coasters, aes(x = Type, fill = `Census Region`)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Number of Roller Coasters by Census Region"
  )

```
